\documentclass[10pt]{article}
\usepackage[margin=0.25in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{multicol}
\usepackage{nicematrix}
\setlength{\columnsep}{0.7cm} 
\setlength{\parindent}{0in}
\title{\raggedright \large Comprehensive Exam Notesheet \hfill Your Name \hfill December 2024 \vspace{-3em}}
\date{}

\begin{document}
\textbf{Comprehensive Exam Notesheet \hfill Your Name \hfill December 2024}
\begin{multicols}{2}

\section*{Chapter One: First Order Equations}
\subsection*{1.3 Separable Equations \& Applications}
\textbf{Key Formula:}
\[
p(y) \frac{dy}{dx} = q(x) \quad \Rightarrow \quad \int \frac{1}{p(y)} \, dy = \int q(x) \, dx
\]
\textbf{Example:} Solve \( \frac{dy}{dx} = 2xy^2 \), \( y(0) = 1 \).
\[
\int y^{-2} \, dy = \int 2x \, dx \quad \Rightarrow \quad -y^{-1} = x^2 + C
\]
Solution: \( y(x) = \frac{-1}{x^2 + C} \).  
\textbf{Application:} Population growth, cooling laws, mixing problems.

\subsection*{1.2 Equilibrium \& Stability}
\textbf{Equilibrium Solutions:} \( f(y) = 0 \) for autonomous equations.  
\textbf{Stability Analysis:} 
\begin{itemize}
    \item \textbf{Stable (sink):} Solutions converge to equilibrium.
    \item \textbf{Unstable (source):} Solutions diverge from equilibrium.
\end{itemize}
\textbf{Example:} \( \frac{dy}{dx} = y^2 - 4y + 3 \).  
Equilibria: \( y = 1, 3 \). Stability: \( y = 1 \) (sink), \( y = 3 \) (source).  

\subsection*{1.4 Linear Equations \& Applications}
\textbf{General Form:} \( \frac{dy}{dx} + P(x)y = Q(x) \).  
\textbf{Integrating Factor:}  
\[
\mu(x) = e^{\int P(x)dx}
\]
\textbf{Example:} Solve \( \frac{dy}{dx} + y = x \).  
\[
\mu(x) = e^x, \quad y = e^{-x}\int xe^x \, dx
\]
Solution: \( y = x - 1 + Ce^{-x} \).

\subsection*{1.5 Bernoulli \& Exact Equations}
\textbf{Bernoulli Equation:} \( \frac{dy}{dx} + P(x)y = Q(x)y^n \).  
\textbf{Transformation:} \( v = y^{1-n} \), reduces to linear.  
\textbf{Example:} Solve \( \frac{dy}{dx} + 2y = 3y^2 \).  
Substitute \( v = y^{-1} \):  
\[
\frac{dv}{dx} - 2v = -3
\]
Solution: \( v = Ce^{-2x} - \frac{3}{2} \), so \( y = \frac{1}{Ce^{-2x} - \frac{3}{2}} \).

\textbf{Exact Equation:}  
\[
M(x, y) + N(x, y)y' = 0 \quad \text{if } \frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}.
\]
\textbf{Example:} Solve \( (2x + y) + (x + 3y)y' = 0 \).  
Check exactness:  
\[
\frac{\partial M}{\partial y} = 1, \quad \frac{\partial N}{\partial x} = 1
\]
Integrate \( M(x, y) \):  
\[
\phi(x, y) = x^2 + xy + \frac{y^2}{2} + C
\]

\section*{Chapter Two: Second Order Equations}
\subsection*{2.1-2.3 General Solutions \& Constant Coefficients}
\textbf{General Form:} \( ay'' + by' + cy = 0 \).  
\textbf{Characteristic Equation:} \( ar^2 + br + c = 0 \).  
\textbf{Case 1:} Distinct Roots \( r_1, r_2 \):  
\[
y(t) = C_1e^{r_1t} + C_2e^{r_2t}
\]
\textbf{Example:} Solve \( y'' - 3y' + 2y = 0 \).  
Roots: \( r_1 = 2, r_2 = 1 \).  
Solution: \( y(t) = C_1e^{2t} + C_2e^t \).

\subsection*{2.4-2.6 Mechanical Vibrations}
\textbf{Free Vibrations:} \( mx'' + kx = 0 \).  
Solution: \( x(t) = C_1\cos(\omega t) + C_2\sin(\omega t) \), \( \omega = \sqrt{\frac{k}{m}} \).  
\textbf{Forced Vibrations:} \( mx'' + cx' + kx = F(t) \).  
Solve using undetermined coefficients or variation of parameters.  
\textbf{Example:} \( x'' + 4x = \sin(2t) \).  
Solution: \( x(t) = C_1\cos(2t) + C_2\sin(2t) + \frac{1}{4}\sin(2t) \).

\section*{Chapter Three: Laplace Transform}
\textbf{Key Properties:}
\[
\mathcal{L}\{f'(t)\} = sF(s) - f(0), \quad \mathcal{L}\{f''(t)\} = s^2F(s) - sf(0) - f'(0)
\]
\textbf{Example:} Solve \( y'' + 2y' + y = e^{-t}, \, y(0) = 0, \, y'(0) = 1 \).  
\[
s^2Y(s) + 2sY(s) + Y(s) = \frac{1}{s+1}
\]
Solution: \( Y(s) = \frac{1}{(s+1)^3} \). Inverse transform: \( y(t) = \frac{t^2}{2}e^{-t} \).

\section*{Chapter Four: Systems of Equations \& Matrices}
\subsection*{Gaussian Elimination \& Rank}
\textbf{Steps:}
\begin{itemize}
    \item Forward elimination to achieve row-echelon form.
    \item Back substitution to solve.
\end{itemize}
\textbf{Example:}
\[
\begin{bNiceMatrix}
1 & 2 & -1 & | & 8 \\ 
0 & 1 & 3 & | & -11 \\ 
0 & 0 & 1 & | & -3
\end{bNiceMatrix}
\]
Solution: \( x_3 = -3, x_2 = -2, x_1 = 4 \).

\section*{Chapter Six: Eigenvalues \& Eigenvectors}
\textbf{Characteristic Polynomial:}
\[
\text{det}(A - \lambda I) = 0
\]
\textbf{Example:}
\[
A = \begin{bNiceMatrix}
4 & 2 \\ 
1 & 3
\end{bNiceMatrix}, \quad \text{det}(A - \lambda I) = (\lambda - 5)(\lambda - 2)
\]
Eigenvalues: \( \lambda = 5, 2 \). Eigenvectors: Solve \( (A - \lambda I)x = 0 \).

\section*{Chapter Seven: Systems of Differential Equations}
\textbf{Eigenvalue Method:} For \( \vec{x}' = A\vec{x} \), solve eigenvalue problem for \( A \).  
\textbf{Example:}
\[
\vec{x}' = \begin{bNiceMatrix}
2 & -1 \\ 
1 & 2
\end{bNiceMatrix}\vec{x}
\]
Eigenvalues: \( \lambda = 3, 1 \). Solution:
\[
\vec{x}(t) = c_1e^{3t}\vec{v_1} + c_2e^{t}\vec{v_2}
\]

\end{multicols}
\end{document}
