\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{lipsum}

\title{CS1800 Homework 2 Solutions}
\author{}
\date{}

% Header for every page
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Name: Sean Balbale}
\fancyhead[R]{HW Group: None}

\begin{document}

\maketitle
\newpage

\section*{Problem 1: English to Logic}

Given:
\begin{itemize}
    \item $a$: The alarm is ringing
    \item $b$: The battery is low
    \item $c$: The system is on
\end{itemize}

\subsection*{i. The alarm is ringing and the system is on}
\[
a \land c
\]

\subsection*{ii. The alarm is not ringing but the system is on}
\[
\neg a \land c
\]

\subsection*{iii. The system is not on and the alarm is not ringing}
\[
\neg c \land \neg a
\]

\subsection*{iv. There is no way that the alarm is ringing}
\[
\neg a
\]

\subsection*{v. The alarm is ringing or the battery is low}
\[
a \lor b
\]

\subsection*{vi. Despite the fact that the alarm is ringing, the battery is low}
\[
a \land b
\]
\newpage

\section*{Problem 2: Arithmetic \& Logic}

\subsection*{i. $(7 \geq 7) \land (8 < 5)$}
\begin{align*}
7 \geq 7 &\text{ is True} \\
8 < 5 &\text{ is False} \\
\text{True} \land \text{False} &= \text{False}
\end{align*}
\[
\text{False}
\]

\subsection*{ii. $(7 \geq 7) \lor (8 < 5)$}
\begin{align*}
7 \geq 7 &\text{ is True} \\
8 < 5 &\text{ is False} \\
\text{True} \lor \text{False} &= \text{True}
\end{align*}
\[
\text{True}
\]

\subsection*{iii. $(6 = 3 + 3) \land (10 < 1)$}
\begin{align*}
6 = 3 + 3 &\text{ is True} \\
10 < 1 &\text{ is False} \\
\text{True} \land \text{False} &= \text{False}
\end{align*}
\[
\text{False}
\]

\subsection*{iv. $\neg(9 = 8)$}
\begin{align*}
9 = 8 &\text{ is False} \\
\neg \text{False} &= \text{True}
\end{align*}
\[
\text{True}
\]

\subsection*{v. $\neg((6 = 3 + 3) \lor (10 < 1))$}
\begin{align*}
6 = 3 + 3 &\text{ is True} \\
10 < 1 &\text{ is False} \\
\text{True} \lor \text{False} &= \text{True} \\
\neg \text{True} &= \text{False}
\end{align*}
\[
\text{False}
\]
\newpage

\section*{Problem 3: Tarski World Predicates \& Quantifiers}

\subsection*{i. $\text{star}(c) \land \neg \text{shade}(c)$}
"c is a star and c is not shaded."
\[
\text{False}
\]

\subsection*{ii. $\exists x (\text{circ}(x) \land \text{shade}(x))$}
"There exists an x such that x is a circle and x is shaded."
\[
\text{True}
\]

\subsection*{iii. $\forall x (\text{square}(x) \rightarrow \neg \text{shade}(x))$}
"For all x, if x is a square, then x is not shaded."
\[
\text{True}
\]

\subsection*{iv. $\forall x \forall y (\text{star}(x) \land \neg \text{shade}(x) \land \text{next to}(x, y)) \rightarrow (\text{shade}(y) \land \text{circ}(y))$}
"For all x and y, if x is a star and not shaded, and x is next to y, then y is shaded and is a circle."
\[
\text{False}
\]

\subsection*{v. $\exists x \forall y (\text{next to}(x, y))$}
"There exists an x such that for all y, x is next to y."
\[
\text{False}
\]

\subsection*{vi. $\forall y \exists x (\text{shade}(x) \land \text{next to}(x, y))$}
"For all y, there exists an x such that x is shaded and x is next to y."
\[
\text{True}
\]
\newpage

\section*{Problem 4: Converse, Inverse, Contrapositive}

Given the statement:
\[
\forall x (\neg \text{shade}(x) \rightarrow \text{circle}(x))
\]

\subsection*{i. True or False?}
The statement is \textbf{True}.

\subsection*{ii. Contrapositive:}
\[
\forall x (\neg \text{circle}(x) \rightarrow \text{shade}(x))
\]
The contrapositive is \textbf{True}.

\subsection*{iii. Converse:}
\[
\forall x (\text{circle}(x) \rightarrow \neg \text{shade}(x))
\]
The converse is \textbf{False}.

\subsection*{iv. Inverse:}
\[
\forall x (\text{shade}(x) \rightarrow \neg \text{circle}(x))
\]
The inverse is \textbf{False}.

\subsection*{v. True conditional statement whose converse is False:}
"If an object is a rectangle, then it is not shaded."
The converse is False if there is an unshaded circle.

\subsection*{vi. True conditional statement whose contrapositive is False:}
This situation is not possible. If the conditional is True, the contrapositive must also be True.
\newpage

\section*{Problem 5: Vending Machine}

Given:
\begin{itemize}
    \item $E$: The machine is empty.
    \item $S$: The user made a selection.
    \item $P$: The user paid a quarter.
    \item $V$: The machine will vend a soda.
    \item $R$: The machine will return the quarter.
\end{itemize}

\subsection*{i. Truth Table for $V$ and $R$:}

\[
\begin{array}{|c|c|c|c|c|}
E & S & P & V & R \\
\hline
0 & 0 & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 \\
0 & 1 & 1 & 1 & 0 \\
1 & 0 & 0 & 0 & 0 \\
1 & 0 & 1 & 0 & 1 \\
1 & 1 & 0 & 0 & 0 \\
1 & 1 & 1 & 0 & 1 \\
\end{array}
\]

\subsection*{ii. Expression for $V$:}
\[
V = \neg E \land S \land P
\]

\subsection*{iii. Expression for $R$:}
\[
R = E \land P
\]
\newpage

\section*{Problem 6: English to Logic (Part 2)}

Given:
\begin{itemize}
    \item $h$: They have hamsters.
    \item $w$: They have whales.
    \item $d$: They have dinosaurs.
\end{itemize}

\subsection*{i. The zoo doesn’t have any hamsters or whales, but it does have dinosaurs.}
\[
\neg h \land \neg w \land d
\]

\subsection*{ii. The zoo has at least two of the three groups of animals.}
\[
(h \land w) \lor (h \land d) \lor (w \land d)
\]

\subsection*{iii. The zoo has exactly two of the three groups of animals.}
\[
(h \land w \land \neg d) \lor (h \land d \land \neg w) \lor (w \land d \land \neg h)
\]

\subsection*{iv. The zoo has, at most, one of these groups of animals.}
\[
(\neg h \land \neg w \land d) \lor (\neg h \land w \land \neg d) \lor (h \land \neg w \land \neg d)
\]

\end{document}
