%==============================================================================
% SEAN BALBALE — CURRICULUM VITAE
%==============================================================================
\documentclass[10pt, letterpaper]{article}
\usepackage[top=0.65in,bottom=0.65in,left=0.70in,right=0.70in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{booktabs}

\hypersetup{colorlinks=true,urlcolor=black,linkcolor=black,
  pdfauthor={Sean Balbale},pdftitle={Sean Balbale --- Curriculum Vitae}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt}

\titleformat{\section}{\normalfont\normalsize\bfseries\uppercase}{}{0em}{}
  [\vspace{-4pt}\rule{\linewidth}{0.6pt}\vspace{2pt}]
\titlespacing*{\section}{0pt}{10pt}{4pt}

\setlist[itemize]{leftmargin=1.2em,itemsep=1pt,topsep=2pt,parsep=0pt,partopsep=0pt}

\newcommand{\cventry}[4]{%
  \begin{tabularx}{\linewidth}{@{}X r@{}}
    \textbf{#1} \textnormal{---} \textit{#2} & \textnormal{#4} \\
    \multicolumn{2}{@{}l@{}}{\small #3}
  \end{tabularx}\vspace{1pt}}

\newcommand{\cventryshort}[3]{%
  \begin{tabularx}{\linewidth}{@{}X r@{}}
    \textbf{#1} \textnormal{---} \textit{#2} & \textnormal{#3}
  \end{tabularx}\vspace{1pt}}

\newcommand{\cvproject}[2]{\textbf{#1} \hfill \textit{#2}\vspace{1pt}}

\begin{document}

%--- HEADER ---
\begin{center}
  {\LARGE \textbf{SEAN BALBALE}} \\[4pt]
  \small
  41 Hillcrest Rd, Weston, MA~02493 \quad\textbar\quad
  +1\,(781)\,290-8862 \quad\textbar\quad
  \href{mailto:sean.balbale@trincoll.edu}{sean.balbale@trincoll.edu} \\[2pt]
  \href{https://www.linkedin.com/in/seanbalbale/}{linkedin.com/in/seanbalbale} \quad\textbar\quad
  \href{https://github.com/sbalbale}{github.com/sbalbale} \quad\textbar\quad
  \href{https://www.seanbalbale.com}{seanbalbale.com}
\end{center}
\vspace{-4pt}

%--- PROFILE ---
%--- PROFILE ---
\section{Profile}

I got into engineering because I wanted to understand how things actually work ---
not just how to use them. That instinct runs through everything I build: bare-metal
Assembly for an 8051 microcontroller, real-time C++ firmware on an Arduino platform,
and a senior capstone built around a fatigue-prediction algorithm running entirely on
a Nordic nRF52840 --- no cloud offload, sub-100\,ms latency. The constraint of the
edge is something I find genuinely satisfying; it forces precision that higher-level
abstractions let you avoid. On the other side of the stack, I owned a Spring Boot~2
to Spring Boot~3 framework migration across multiple production Java services at
Bullhorn, and this summer I am joining Sonos to build the backend microservices
powering user identity and subscriptions across the platform. The problems are
different at each layer --- but the methodology isn't.\\


Beyond implementation, I research the systems we are building toward.
``Embodied Harms and Inferred Data: Redefining Privacy in Extended Reality''
argues that existing legal frameworks are structurally inadequate for immersive
environments that passively extract behavioral biometrics, and proposes a
``Bodyright'' framework that protects human autonomy by design rather than by policy.
It won the Alumni Prize in English Composition at Trinity's 2026 Honors Day and is
forthcoming in \textit{The International Journal of Law, Ethics, and Technology}.
Outside the lab, I compete on Trinity's varsity rowing team and serve as Technology
Chair for the Accidentals. One is about power and synchronization at 5\,a.m.\ on the
Connecticut River; the other is about blending twelve voices into something that sounds
like one. The underlying discipline is identical: precise timing, collective
execution, and zero tolerance for drift.

%--- EDUCATION ---
\section{Education}

\begin{tabularx}{\linewidth}{@{}X r@{}}
  \textbf{Trinity College} \textnormal{---} Hartford, CT & \textit{Expected May 2027}
\end{tabularx}
\vspace{1pt}
\textit{Bachelor of Science in Electrical and Computer Engineering} \\
\textit{Bachelor of Science in Computer Science} \\
\textbf{GPA:} 3.65 / 4.00

\vspace{4pt}
\textbf{Coursework Completed:} Data Structures \& Algorithms, Embedded Systems Design,
Basic Economic Principles, Age of Augustus, Ancient Greek Philosophy, Microelectronic Circuits,
Applied Linear Algebra, Calculus III, Intro to Computer Systems, Writing for a Digital World.\\[2pt]
\textbf{In Progress (Spring 2026):} High-Performance Computing, Deep Learning, Software
Engineering, International Trade, Measurements, Instrumentation \& Analysis, Engineering
Materials, Linear Circuit Theory (ENGR~212L).\\[2pt]
\textbf{Planned (Fall 2026):} Automatic Control Systems, Mechanics I, Computer Science Seminar,
Capstone Design I --- CPSC\,498 / ECE\,483.\\[2pt]
\textbf{AP Credit Transferred:} Computer Science A, English Composition \& Literature,
Physics C: Mechanics, Physics C: Electricity \& Magnetism.

\vspace{4pt}
\textbf{Extracurriculars:} Varsity Athlete --- Trinity College Men's Rowing (Division III);
Singer \& Technology Chair --- The Trinity College Accidentals (premier a cappella ensemble).

\vspace{6pt}

\begin{tabularx}{\linewidth}{@{}X r@{}}
  \textbf{Carnegie Mellon University} \textnormal{---} Pittsburgh, PA (Summer Program) &
  \textit{June -- August 2022}
\end{tabularx}
\vspace{1pt}
Completed university-level coursework: Fundamentals of Programming \& Computer Science (15-112);
Reasoning with Data.

\vspace{6pt}

\begin{tabularx}{\linewidth}{@{}X r@{}}
  \textbf{Northeastern University} \textnormal{---} Boston, MA \& London, UK &
  \textit{September 2023 -- December 2024}
\end{tabularx}
\vspace{1pt}
Attended Northeastern University for three semesters before transferring to Trinity College
(Spring 2025). All coursework transferred.

\vspace{2pt}
\textit{Freshman Fall 2023 --- Boston, MA:}
General Chemistry I, Cornerstone of Engineering 1, Calculus I for Science \& Engineering,
Musical Communities of Boston.

\vspace{2pt}
\textit{Freshman Spring 2024 --- London, UK (Study Abroad):}
Fundamentals of Computer Science II (with Lab), Cornerstone of Engineering 2,
Calculus II for Science \& Engineering, British Drama/London Stage.

\vspace{2pt}
\textit{Sophomore Fall 2024 --- Boston, MA:}
Discrete Structures (with Seminar), Circuits/Signals: Biomedical Applications,
Embedded Design: Enabling Robotics, Differential Equations \& Linear Algebra for Engineering,
First-Year Seminar, Chorus.

\vspace{6pt}

\begin{tabularx}{\linewidth}{@{}X r@{}}
  \textbf{Weston High School} \textnormal{---} Weston, MA & \textit{September 2019 -- June 2023}
\end{tabularx}
\vspace{1pt}
\textit{High School Diploma} \quad \textbf{GPA:} 3.85 / 4.00

\vspace{2pt}
\textbf{Activities:} Varsity Captain --- Wayland-Weston Men's Rowing;
Varsity Ski Team; Concert Choir; Jazz Choir.

%--- SKILLS ---
\section{Technical Skills}

\begin{tabularx}{\linewidth}{@{} l X @{}}
  \textbf{Languages} &
    Java, Python, C, C++, TypeScript, JavaScript (Node.js), Assembly (8051, MIPS), Verilog,
    SQL, MATLAB, \LaTeX \\[3pt]
  \textbf{Backend} &
    Spring Boot 3, Jakarta EE, JUnit 5, Mockito, Node.js, REST APIs \\[3pt]
  \textbf{Frontend / Full-Stack} &
    React, Next.js, Tailwind CSS, AngularJS \\[3pt]
  \textbf{Cloud \& DevOps} &
    AWS, Kubernetes, Docker, Linux, Git (feature branches, merge requests, code review),
    CI/CD pipelines, Log4j2, Slurm, MPI \\[3pt]
  \textbf{Embedded / Hardware} &
    Arduino, ESP32, 8051 Microcontroller, Intel Cyclone~V FPGA, Nordic nRF52840,
    SPI/I\textsuperscript{2}C, BLE~5, ANT+, FreeRTOS, NI-DAQ, Oscilloscopes, Logic Analyzers,
    LTspice \\[3pt]
  \textbf{Mechanical / CAD} &
    SolidWorks, FDM 3D Printing (Klipper firmware), Laser Cutters, Precision Metrology
    Calibration \\[3pt]
  \textbf{Other} &
    Systems Engineering, Zero-Knowledge Architectures, Object-Oriented Design,
    Parallel Computing, MCP (Model Context Protocol) Server Development
\end{tabularx}

%--- WORK EXPERIENCE ---
\section{Work Experience}

\cventry{Cloud/Service Software Engineering Intern}{Sonos}{Boston, MA}{Summer 2026}
\begin{itemize}
  \item Joining the Software Services Cloud Accounts team to architect and maintain backend Java
        microservices powering core user-identity, subscription entitlement, player-settings,
        and API-permission systems across the Sonos platform.
  \item Working within an AWS and Kubernetes environment on containerized cloud services that
        underpin the user experience for all Sonos hardware products globally.
  \item Applying containerization, orchestration, and source-control best practices (GitHub) as
        specified in team pre-internship preparation guidelines.
\end{itemize}

\vspace{6pt}

\cventry{Engineering Teaching Assistant --- ENGR\,212L}{Trinity College}{Hartford, CT}{Spring 2026}
\begin{itemize}
  \item Instructed students through weekly laboratory sessions for Linear Circuit Theory
        (ENGR~212L), covering node and mesh analysis, natural and forced responses of first- and
        second-order RLC circuits, operational amplifiers, and passive and active filter design.
  \item Guided students through the application of Laplace transforms and Fourier series as
        analytical tools for solving linear circuit equations, reinforcing lecture content with
        hands-on hardware experimentation.
  \item Delivered objective data analysis feedback on student lab write-ups to reinforce core
        ECE measurement and instrumentation practices.
\end{itemize}

\vspace{6pt}

\cventry{Software Engineering Intern}{Bullhorn}{Boston, MA}{Summer 2025}
\begin{itemize}
  \item Contributed to enterprise-scale backend Java services across the Shield and Builders
        teams, supporting the SEARCH\_AI, ATS\_CRM, and MIDDLE\_OFFICE engineering groups.
  \item Owned a major framework migration (BH Boot 1.0.0): upgraded multiple production Java
        applications from Spring Boot~2 to Spring Boot~3, systematically refactoring all
        \texttt{javax} namespaces to \texttt{jakarta} and restructuring the proprietary data
        access layer.
  \item Led a critical security initiative to upgrade core logging infrastructure to Log4j2,
        hardening the dependency chain across affected services.
  \item Refactored legacy test suites from TestNG/JMockit to JUnit~5 and Mockito, increasing
        test coverage and long-term maintainability.
  \item Participated in the full software development lifecycle: managed feature branches,
        authored merge requests, and conducted peer code reviews.
\end{itemize}

\vspace{6pt}

\cventry{Metrology Intern}{East Coast Metrology --- A Division of In-Place Machining Company}{Topsfield, MA}{Summer 2024}
\begin{itemize}
  \item Conducted in-house and on-site calibrations for high-performance precision measurement
        equipment, including portable CMMs, total stations, large-volume laser scanners, and
        laser trackers.
  \item Measured manufactured engineering parts against client SolidWorks CAD models to verify
        dimensional compliance with specified tolerances.
  \item Updated and maintained internal-use SolidWorks CAD models to reflect as-calibrated
        equipment configurations.
\end{itemize}

\vspace{6pt}

\cventry{Engineering Intern}{Titan Advanced Energy Solutions}{Salem, MA}{Summer 2023}
\begin{itemize}
  \item Designed and implemented Python software to parse raw JSON output from ultrasound
        sensors and render waveform visualizations, providing a complementary view to existing
        heat-map outputs for non-destructive evaluation of electric vehicle battery packs.
  \item Visualization tooling directly accelerated the detection and classification of physical
        defects within battery cells, supporting quality assurance workflows.
\end{itemize}

\vspace{6pt}

\cventry{Apprentice Optician}{Warby Parker}{Boston, MA}{June 2023 -- January 2024}
\begin{itemize}
  \item Validated prescriptions, took optical measurements, and sold eyewear in a high-volume
        retail environment; worked part-time (5 hours per week) during the academic year and
        full-time during summer 2023.
  \item Ranked highest in sales for the Northeast region for three of the eight weeks worked
        full-time during the summer.
  \item Resolved customer complaints and maintained a high-quality client experience under
        pressure; developed skills in sales, customer service, and teamwork.
\end{itemize}

\vspace{6pt}

\cventry{Founder \& IT Consultant}{Cherrybrook Networks}{Weston, MA}{March 2019 -- May 2025}
\begin{itemize}
  \item Founded and operated a freelance software development and technical support business
        serving local companies, students, and families in the greater Weston, MA area.
  \item Installed and configured software applications; designed and assembled custom computers
        tailored to client specifications.
  \item Deployed residential and small-business Wi-Fi infrastructure; automated secure cloud
        backup systems for client data protection.
  \item Maintained long-term client relationships over six years, developing skills in client
        communication, project scoping, and independent problem-solving.
\end{itemize}

\vspace{6pt}

\cventry{Engineering Teaching Assistant}{Weston Public Schools}{Weston, MA}{October 2022 -- May 2023}
\begin{itemize}
  \item Developed and delivered course curriculum for high school engineering and physics
        courses; wrote and graded assessments and provided individualized academic support.
  \item Maintained and calibrated precision workshop equipment, including FDM 3D printers,
        laser cutters, band saws, and woodworking tools.
\end{itemize}

\vspace{6pt}

\cventry{Rowing Coach}{Wayland-Weston Rowing Association}{Wayland, MA}{September 2019 -- June 2023}
\begin{itemize}
  \item Coached middle school athletes in on-water rowing technique during spring and fall crew
        seasons, building a foundation in catch timing, blade work, and race-pace conditioning.
  \item Built team culture and improved athletes' physical and mental fitness across four
        competitive seasons.
\end{itemize}

%--- RESEARCH ---
\section{Research \& Publications}

\textbf{``Embodied Harms and Inferred Data: Redefining Privacy in Extended Reality''} \\
\textit{The International Journal of Law, Ethics, and Technology}
\hfill \textit{In Production --- Expected Late May 2026}

\vspace{3pt}
\begin{itemize}
  \item Authored a peer-reviewed synthesis paper examining the political economy of Extended
        Reality (XR) technology and its implications for privacy law and human autonomy.
  \item Argued that existing legislative frameworks (BIPA, GDPR) are structurally inadequate
        for immersive environments that passively collect behavioral biometrics --- including
        physiological signals, gaze patterns, and locomotion data --- without active user
        consent.
  \item Deconstructed the ``inference gap'' --- the mechanism by which immersive architectures
        bypass cognitive defenses to extract sensitive psychometric traits (health status,
        political orientation, sexual orientation) without user awareness.
  \item Conducted a comparative analysis of the Apple--Meta dichotomy, demonstrating that
        market-led ``privacy-by-design'' functions primarily as an anti-competitive moat,
        producing a ``Walled Garden'' paradox in which privacy becomes a luxury good.
  \item Applied the legal doctrines of ``undue influence'' and ``defect of consent'' to the
        ``Proteus Effect'' to establish that XR environments possess a unique capacity to
        vitiate user autonomy through subliminal behavioral modification.
  \item Proposed a transition from the ``notice-and-choice'' legal paradigm to a ``Bodyright''
        framework, reclassifying biometric data as an inalienable extension of the human
        sovereign rather than an extractable raw material.
\end{itemize}

%--- HONORS ---
\section{Honors \& Awards}

\cventryshort{Alumni Prize in English Composition}{Trinity College --- Department of English}{May 2026}
\begin{itemize}
  \item Received the Alumni Prize in English Composition at the Trinity College Honors Day
        Ceremony (May 1, 2026), awarded by the Department of English faculty to recognize the
        best expository writing produced in the year.
  \item Prize awarded for ``Embodied Harms and Inferred Data: Redefining Privacy in Extended
        Reality,'' accepted for publication in \textit{The International Journal of Law,
        Ethics, and Technology} (Expected Late May 2026).
\end{itemize}

%--- PROJECTS ---
\section{Selected Projects}

\cvproject{Edge-Compute Pacing \& Feedback Node (Senior Capstone)}{Fall 2026 -- Spring 2027}
\begin{itemize}
  \item An embedded edge-compute node that delivers real-time biomechanical and fatigue
        feedback to endurance athletes (rowers and cyclists) without cloud or smartphone offload.
  \item Hardware platform: Nordic nRF52840 (ARM Cortex-M4F @ 64\,MHz, 256\,kB RAM, 1\,MB Flash)
        with native dual-mode BLE~5 and ANT+ radio; WS2812B NeoPixel LED strip and piezo
        buzzer as output peripherals; 18650 Li-Ion cell with BMS for $\geq$6\,h runtime.
  \item Core algorithm: discrete-time W$'$ balance model (Skiba et al., 2012) implemented as a
        first-order Euler update in C/C++ on FreeRTOS, targeting $\leq$100\,ms end-to-end
        sensor-to-LED latency on a bare-metal microcontroller.
  \item Ingests live power and heart-rate data via ANT+ Bicycle Power Profile (PWR) and Heart
        Rate Profile (HRM); BLE GATT power service as fallback.
  \item Open research questions: (1) algorithmic correctness under Euler discretisation vs.\
        continuous ground truth; (2) optimal FreeRTOS scheduling policy for worst-case latency
        on a single-core MCU; (3) feasibility of a lightweight LSTM for personalised
        $\tau_{W'}$ estimation beyond population-level constants.
\end{itemize}

\vspace{4pt}

\cvproject{Personal Portfolio Website}{Spring 2026}
\begin{itemize}
  \item Architected and deployed a production full-stack web application at
        \href{https://www.seanbalbale.com}{seanbalbale.com}
        using Next.js, React, and TypeScript; styled with Tailwind CSS and modern UI component
        libraries.
  \item Built a custom backend API route to handle contact form submissions and programmatically
        dispatch transactional emails; implemented dark/light mode theming and an embedded PDF
        resume viewer with dynamic project routing.
  \item Deployed via Vercel with CI/CD on each Git push.
\end{itemize}

\vspace{4pt}

\cvproject{Remote MCP Server (Obsidian Vault Integration)}{Spring 2026}
\begin{itemize}
  \item Built a TypeScript MCP (Model Context Protocol) server that exposes a personal Obsidian
        knowledge vault over a secure remote endpoint.
  \item Implemented Bearer token authentication and deployed via a Cloudflare Tunnel / Tailscale
        network overlay; orchestrated with Docker Compose multi-profile deployment supporting
        multiple environment configurations.
\end{itemize}

\vspace{4pt}

\cvproject{HPC Cluster --- High-Performance Computing Lab}{Fall 2025}
\begin{itemize}
  \item Assembled and configured a personal three-node HPC cluster using Lenovo ThinkStation
        P340~SSF machines interconnected via a dedicated Gigabit Ethernet switch.
  \item Installed and configured Slurm workload manager for job scheduling; benchmarked parallel
        workloads using MPI for CPSC\,375 (High-Performance Computing).
\end{itemize}

\vspace{4pt}

\cvproject{Project Aletheia --- Interactive XR Privacy Manifesto}{Fall 2025}
\begin{itemize}
  \item Built a full-stack React web application serving as an interactive privacy manifesto,
        demonstrating concrete biometric tracking risks inherent in Extended Reality environments.
  \item Advocates for Privacy-by-Design principles and client-side Zero-Knowledge architectures
        as a technical implementation of the ``Bodyright'' framework developed in published
        research.
\end{itemize}

\vspace{4pt}

\cvproject{Embedded Heart Rate Monitor (8051 Assembly)}{Spring 2025}
\begin{itemize}
  \item Designed and wired the complete hardware system for a clinical heart rate monitor and
        mechanical click counter on a custom breadboard.
  \item Wrote low-level Assembly language firmware for an 8051 microcontroller handling external
        analog input reading, interrupt-driven signal processing, ADC sampling, and real-time BCD
        display output --- entirely without a C runtime or operating system.
\end{itemize}

\vspace{4pt}

\cvproject{Haptic Horizon --- Wearable Navigation Device for the Visually Impaired}{Spring 2024}
\begin{itemize}
  \item Designed and built a wearable haptic alarm system to assist visually impaired users in
        navigating physical environments.
  \item Wrote embedded C++ firmware for an Arduino-based platform integrating an HC-SR04
        ultrasonic ranging sensor and a haptic motor; sensor output encoded distance and direction
        into distinct vibration patterns without any auditory output.
  \item Handled interrupt-driven sensor polling, signal conditioning, and real-time actuator
        control within a single-threaded embedded event loop.
\end{itemize}

\vspace{4pt}

\cvproject{Robotic Control via FPGA}{Fall 2024}
\begin{itemize}
  \item Implemented a Verilog hardware description on an Intel Cyclone~V FPGA to control a
        robotic system, demonstrating direct hardware-software co-design principles.
  \item Designed logic modules for motor control signals, sensor interfacing, and real-time
        feedback within the timing and resource constraints of the target FPGA fabric.
\end{itemize}

\vspace{4pt}

\cvproject{Content-Aware Java Image Editor --- Seam Carving (CS2510 Term Project)}{Spring 2024}
\begin{itemize}
  \item Implemented the seam carving algorithm (Avidan \& Shamir, 2007) for content-aware image
        resizing in Java, dynamically identifying and removing low-energy pixel seams to resize
        images while preserving salient visual content.
  \item Built an interactive command-line menu supporting multiple editing operations; applied
        object-oriented design patterns throughout using Java and Apache Maven.
\end{itemize}

\vspace{4pt}

\cvproject{``Fail Early and Often'' --- Python Roguelite (CMU 15-112 Term Project)}{Summer 2022}
\begin{itemize}
  \item Designed and implemented a fully-featured roguelite RPG in Python as the term project
        for CMU CS~15-112 (Fundamentals of Programming \& CS); approximately 25 hours of
        development.
  \item Built core game systems: procedural map generation, WASD movement, proximity-based melee
        combat, a multi-slot inventory system with consumable potions, enemy AI wave logic, sprite
        rendering, collision detection, and a toggle-able cone-of-vision mechanic.
  \item Demo: \href{https://youtu.be/aftX408l35E}{youtu.be/aftX408l35E}
\end{itemize}

\vspace{4pt}

\cvproject{Space Invaders --- Python Arcade Game}{2020}
\begin{itemize}
  \item Designed and implemented a Space Invaders clone in Python using Pygame, practicing sprite
        rendering, collision detection, enemy AI wave logic, and a high-score persistence system.
\end{itemize}

%--- ACTIVITIES ---
\section{Activities \& Interests}

\textbf{Athletics:} Varsity Rowing, Trinity College Men's Crew (Division~III). Outside of crew, training interests include triathlon,
road and gravel cycling (power meter training, pedal-stroke symmetry analysis), ski mountaineering,
technical mountaineering, and backcountry backpacking.

\vspace{4pt}
\textbf{Cooking:} Serious home cook across a wide range of cuisines --- tacos, samosas, katsu
ramen, Indian, Italian, and whatever fusion experiment sounds right that week. The kitchen is
where precision and improvisation meet, which is probably why it keeps me coming back.

\vspace{4pt}
\textbf{Music \& Technology:} Technology Chair and Singer, Trinity College Accidentals (premier
a cappella group) --- manage live sound engineering, recording infrastructure, and digital
production for performances and studio sessions.

\vspace{4pt}
\textbf{Engineering \& Making:} 3D printing on a heavily modified Creality Ender~3~v2 running
Klipper open-source firmware.

\vspace{4pt}
\textbf{Research Interests:} Privacy law (GDPR, BIPA, biometric autonomy); XR ethics and
surveillance capitalism; hardware-constrained software design; distributed systems and cloud
microservices; HPC and parallel computing; embedded systems for sports science.

%--- REFERENCES ---
\section{References}

Available upon request.

\end{document}