%-------------------------
% ATS-Optimized One-Page Resume — Sean Balbale
%-------------------------
\documentclass[letterpaper,9.8pt]{article}

\usepackage[left=0.5in,right=0.5in,top=0.4in,bottom=0.4in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage[hidelinks]{hyperref}
\usepackage[T1]{fontenc}
\usepackage{xcolor}

\pagestyle{empty}
\raggedbottom
\raggedright

% Section formatting
\titleformat{\section}{\small\bfseries\uppercase}{}{0em}{}[\titlerule]
\titlespacing{\section}{0pt}{3.5pt}{2.5pt}

% Tight lists
\setlist[itemize]{leftmargin=*,topsep=0pt,itemsep=0pt,parsep=0pt,partopsep=0pt}
\renewcommand{\labelitemi}{$\bullet$}

\newcommand{\vsm}{\vspace{1.5pt}}

\begin{document}

%----------HEADER----------
\begin{center}
  {\large\bfseries SEAN BALBALE}\\[1.5pt]
  \footnotesize
  Weston, MA 02493 \;|\;
  +1\,(781)\,290-8862 \;|\;
  \href{mailto:sean.balbale@trincoll.edu}{sean.balbale@trincoll.edu} \;|\;
  \href{https://linkedin.com/in/seanbalbale}{linkedin.com/in/seanbalbale} \;|\;
  \href{https://seanbalbale.com}{seanbalbale.com}
\end{center}
\vspace{-3pt}

%----------EDUCATION----------
\section{Education}

\noindent\textbf{Trinity College} --- Hartford, CT \hfill \textit{Expected May 2027}\\
\small{B.S.\ Electrical \& Computer Engineering \;|\; B.S.\ Computer Science \;|\; GPA: 3.65 / 4.00}\\
\footnotesize{Coursework: High-Performance Computing, Deep Learning, Software Engineering, Embedded Systems Design, Microelectronic Circuits, Data Structures \& Algorithms, Intro to Computer Systems, Applied Linear Algebra, Calculus III.}\\
\footnotesize{Extracurriculars: Varsity Athlete --- Men's Rowing (Div.\ III); Technology Chair \& Singer --- The Trinity College Accidentals.}

\vsm
\noindent\footnotesize{\textbf{Northeastern University} --- Boston, MA \& London, UK \hfill Attended Sept.\ 2023\,--\,Dec.\ 2024; all credits transferred to Trinity}

%----------SKILLS----------
\section{Technical Skills}

\begin{tabular}{@{}l@{\hspace{5pt}}l}
\footnotesize\textbf{Languages:}      & \footnotesize Java, Python, C, C++, TypeScript, Assembly (8051), Verilog, SQL, MATLAB, LaTeX \\
\footnotesize\textbf{Backend/Cloud:}  & \footnotesize Spring Boot 3, Jakarta EE, JUnit 5, Mockito, REST APIs, AWS, Kubernetes, Docker, CI/CD, Log4j2 \\
\footnotesize\textbf{Frontend:}       & \footnotesize React, Next.js, Tailwind CSS, Node.js, AngularJS \\
\footnotesize\textbf{Embedded/HW:}   & \footnotesize Arduino, 8051 MCU, nRF52840 (Cortex-M4F), Intel Cyclone V FPGA, FreeRTOS, BLE 5, ANT+, LTspice \\
\footnotesize\textbf{Tools:}          & \footnotesize Git, Linux, SolidWorks, 3D Printing (Klipper), Slurm/MPI, Precision Metrology, Oscilloscopes \\
\end{tabular}

%----------EXPERIENCE----------
\section{Work Experience}

\noindent\textbf{Sonos} --- Boston, MA \hfill \textit{Summer 2026}\\
\textit{\small Software Engineering Intern --- Cloud Accounts (SW Services)}
\begin{itemize}
  \item\footnotesize Architecting and maintaining backend Java microservices powering user-identity, subscription entitlement, player-settings, and API-permission systems across the Sonos platform.
  \item\footnotesize Working in an AWS and Kubernetes environment; applying containerization, orchestration, and GitHub source-control best practices on cloud services used globally.
\end{itemize}

\vsm
\noindent\textbf{Bullhorn} --- Boston, MA \hfill \textit{Summer 2025}\\
\textit{\small Software Engineering Intern}
\begin{itemize}
  \item\footnotesize Owned framework migration (BH Boot 1.0.0): upgraded multiple production Java applications from Spring Boot 2 to Spring Boot 3, refactoring all \texttt{javax} namespaces to \texttt{jakarta} and restructuring the proprietary data access layer.
  \item\footnotesize Led a critical security initiative upgrading core logging infrastructure to Log4j2; migrated legacy test suites from TestNG/JMockit to JUnit 5 and Mockito across affected services.
  \item\footnotesize Participated in the full SDLC --- managed feature branches, authored merge requests, and conducted peer code reviews (SEARCH\_AI, ATS\_CRM, MIDDLE\_OFFICE teams).
\end{itemize}

\vsm
\noindent\textbf{East Coast Metrology --- A Division of In-Place Machining Co.} --- Topsfield, MA \hfill \textit{Summer 2024}\\
\textit{\small Metrology Intern}
\begin{itemize}
  \item\footnotesize Calibrated portable CMMs, laser trackers, total stations, and large-volume scanners; verified manufactured parts against SolidWorks CAD models for dimensional compliance.
\end{itemize}

\vsm
\noindent\textbf{Titan Advanced Energy Solutions} --- Salem, MA \hfill \textit{Summer 2023}\\
\textit{\small Engineering Intern}
\begin{itemize}
  \item\footnotesize Built Python tooling to parse ultrasound JSON data and render waveform visualizations for non-destructive EV battery defect detection, accelerating QA workflows.
\end{itemize}

\vsm
\noindent\textbf{Cherrybrook Networks (Founder)} --- Weston, MA \hfill \textit{Mar.\ 2019 -- May 2025}\\
\textit{\small Freelance Software Developer \& IT Consultant}
\begin{itemize}
  \item\footnotesize Operated a six-year freelance business delivering software dev, custom PC builds, Wi-Fi infrastructure, and automated cloud-backup solutions to local clients.
\end{itemize}

%----------PROJECTS----------
\section{Selected Projects}

\noindent\textbf{Edge-Compute Pacing \& Feedback Node} \textit{(Senior Capstone --- CPSC 498 / ECE 483)} \hfill \textit{Fall 2026\,--\,Spring 2027}
\begin{itemize}
  \item\footnotesize Embedded edge node providing real-time fatigue feedback for rowers/cyclists via the W$'$ balance model in C/C++ on FreeRTOS on a Nordic nRF52840 (Cortex-M4F, 64 MHz); targets $\leq$100 ms sensor-to-LED latency entirely offline via BLE 5 / ANT+.
\end{itemize}

\vsm
\noindent\textbf{Cherrybrook Networks --- Full-Stack Portfolio} \hfill \textit{2024\,--\,Present}\\
\footnotesize Production Next.js / React / TypeScript web app (Vercel CI/CD); custom backend API for transactional email; 151+ commits, 95.5\% TypeScript.

\vsm
\noindent\textbf{Haptic Horizon --- Wearable Navigation Device} \hfill \textit{2025}\\
\footnotesize Embedded C++ firmware (Arduino): HC-SR04 ultrasonic sensor + haptic motor encodes proximity as vibration patterns for visually impaired users; interrupt-driven real-time event loop.

%----------HONORS----------
\section{Honors \& Research}

\noindent\textbf{Alumni Prize in English Composition} --- Trinity College \hfill \textit{May 2026}\\
\footnotesize{\textit{``Embodied Harms and Inferred Data: Redefining Privacy in Extended Reality''} --- accepted for publication in \textit{The International Journal of Law, Ethics, and Technology} (Expected Late May 2026). Proposes a ``Bodyright'' framework for biometric data autonomy in immersive XR environments; awarded by the Trinity Department of English at Honors Day.}

%----------PERSONAL----------
\section{Personal}

\footnotesize{Competitive varsity rower (Heavyweight eight/four, 2K erg sprints); avid cook --- samosas and katsu ramen; weightlifting, running, ski mountaineering. Manage live sound and recording for a collegiate a cappella ensemble.}

\end{document}