\documentclass[10pt]{article}
\usepackage[margin=0.25in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{multicol}
\usepackage{nicematrix}
\setlength{\columnsep}{0.7cm} 
\setlength{\parindent}{0in}
\title{\raggedright \large Math 2341 Final Exam Notesheet \hfill Sean Balbale \hfill December 2024 \vspace{-3em}}
\date{}

\begin{document}
% \maketitle
\textbf{Math 2341 Final Exam Notesheet \hfill Sean Balbale \hfill December 2024}
\begin{multicols}{2}
\section*{Mixing Problems (1.4)}
\textbf{Differential Equation Form:}
\[
\frac{dA}{dt} = \text{rate in} - \text{rate out}
\]
\textbf{Example 1:}  
Tank with 200 gallons, 50 lbs of salt initially. Brine with \(1\) lb/gal flows in at \(5\) gal/min, same outflow rate.
\[
\frac{dA}{dt} = 5(1) - \frac{5A}{200}, \quad A(t) = 200 - 150e^{-t/40}
\]

\textbf{Example 2:}  
Tank starts with 500 L water and 5 kg salt. Brine with \(0.5\) kg/L enters at \(10\) L/min, \(5\) L/min outflow.
\[
\frac{dA}{dt} = 10(0.5) - \frac{5A}{500+t}
\]

\section*{Exact Equations (1.5)}
\textbf{Form:} \( M(x, y) + N(x, y)y' = 0 \)  
\textbf{Test for Exactness:}  
\[
\frac{\partial M}{\partial y} = \frac{\partial N}{\partial x}
\]
\textbf{Example:}  
\[
(2x + y^2) + (2y + xy)y' = 0
\]
Integrate \( M(x, y) \):  
\[
\phi(x, y) = x^2 + xy^2 + g(y)
\]
Substitute and solve: \( g(y) = C \).

\section*{Laplace Transform (3.2-3.4)}
\textbf{Laplace Transform of Derivatives:}
\[
\mathcal{L}\{f'(t)\} = sF(s) - f(0), \quad \mathcal{L}\{f''(t)\} = s^2F(s) - sf(0) - f'(0)
\]
\textbf{Example: Solve \( y'' + 2y' + y = e^{-t} \), \( y(0) = 0, y'(0) = 1 \).}  
\[
s^2Y(s) + 2sY(s) + Y(s) = \frac{1}{s+1}
\]
Solution:
\[
Y(s) = \frac{1}{(s+1)^2(s+1)}
\]

\section*{Undetermined Coefficients (2.5)}
\textbf{Form:} \( y'' + ay' + by = f(t) \)
where
\begin{enumerate}
    \item f(x) = a polynomial in x
    \item f(x) = \( e^{ax} \) (a polynomial in x)
    \item f(x) = \( e^{ax} \) (cos(bx) or sin(bx)) (a polynomial in x)
\end{enumerate}
then the particular solution $y_p$ is:
\begin{enumerate}
    \item $y_p = x^k$ (a general polynomial of the same degree) \\
    where $k$ is the number of times that 0 is a root of the characteristic equation.
    \item $y_p = x^k e^{ax}$ (a general polynomial of the same degree) \\
    where $k$ is the number of times that $a$ is a root of the characteristic equation.
    \item $y_p = x^k e^{ax}$ [(a general polynomial of the same degree) cos(bx) + (a general polynomial of the same degree) sin(bx)] \\
    where $k$ is the number of times that $a + bi$ is a root of the characteristic equation.
\end{enumerate}

\textbf{Example 1: Solve \( y'' + 3y' + 2y = 4e^{-3x} \).}  
\[
\text{Characteristic roots: } r = -1, -2
\]
Guess \( y_p = x^0 e^{-3x}(A)\), substitute to find \( A \). \\
\textbf{Example 2: Solve \( y'' + 4y = 2cos(3x) \).}
\[
\text{Characteristic roots: } r= 0 \pm 2i
\]
Guess \( y_p = x^0 e^{0t} (A\cos(3t) + B\sin(3t)) \), substitute to find \( A, B \).

\section*{Spring Problems (2.4, 2.6)}
\textbf{Form:} \( mx'' + cx' + kx = F(t) \)  
Cases:
\begin{itemize}
    \item Free Oscillation: \( c = 0, F(t) = 0 \)
    \[
    x(t) = C_1\cos(\omega t) + C_2\sin(\omega t)
    \]
    \item Damped Oscillation (\( c > 0 \)): Analyze overdamped, underdamped, or critically damped cases.
\end{itemize}

\section*{Eigenvalues and Eigenvectors (6.1)}
\textbf{Characteristic Polynomial:}
\[
\text{det}(A - \lambda I) = 0
\]
\textbf{Example:}
\[
A = \begin{bNiceMatrix}
4 & 2 \\ 
1 & 3
\end{bNiceMatrix}, \quad \text{det}(A - \lambda I) = (\lambda - 5)(\lambda - 2)
\]
Eigenvalues: \( \lambda = 5, 2 \).

\section*{Solving Systems of Equations (7.3)}
\textbf{Example:}
\[
\vec{x}' = \begin{bNiceMatrix}
2 & -1 \\ 
1 & 2
\end{bNiceMatrix} \vec{x}
\]
Eigenvalues: \( \lambda = 3, 1 \).  
Solutions:
\[
\vec{x}(t) = c_1e^{3t}\vec{v_1} + c_2e^{t}\vec{v_2}
\]

\end{multicols}
\end{document}
